Probability of causal contact between interstellar civilizations through MonteCarlo simulations

### 2020-05-13 submit_v03/manuscript.tex
pro, writing

%-------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%-------------------------------------------------------------------

In our simulations we also assume the simplest situations for the
growth of the sphere of first contact.
%
For example, we discard the possibility of stellar colonization
\citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
starling_virulence_2013, barlow_galactic_2012, jeong_large_2000,
maccone_mathematical_2011} and assume that the communication aim is
performed to all directions with the same power and the same
probability, i.e., strongly assume isotropic communication in all
cases.
%
More detailed simulations could be produced considering different
efficiency of communication or detection methods.
%
However, in this case we also would complicate the experimental setup,
make the results less evident.
%
The results are independent of the nature of life (organic or
artificial).
%
The lifetime of a civilization can be caused by auto destruction or by
external factors.
%}}}

@@

The temporal structure of the process is defined by two distribution
parameters that represent the mean time interval that a \ceti{}
can emit and receive signals, and the mean time interval
between the emergence of consecutive \cetis{}.
%
The spatial structure of the simulation is given by the size and shape
of the Galactic Habitable Zone and the maximum distance a signal can
travel to be detected (D$_{max}$).
%
The parameters for the temporal distributions also determine the
spatial properties, since the density of active \cetis{} in the galaxy
depend on these two parameters.
%
For example, a small $\tau_a$ and a large $\tau_s$ will produce a
galaxy densely populated with \cetis{}.
%
Also, some hypotheses regarding the shapes of the distributions of the
temporal parameters must be made in order to complete the
simulation.
%
\citet{forgan_spatiotemporal_2011} argues that the times at which
different civilizations become intelligent follow a Gaussian
distribution, and then the distribution of inter-arrival times is an
inverse exponential.
%
We choose to assume that the distribution of the times of A events is a
stationary Poisson process, and then the distribution of the times
between the appearance of new consecutive \cetis{} is exponential.
%
Regarding the duration of a \ceti{}, we propose that the distribution
of the duration of \cetis{} is a stationary exponential distribution.
%
\citet{balbi_impact_2018} stress the fact that it would be desirable
to arrive at a theoretical statistical distribution of the lifespan of
civilizations, preferentially on the basis of the underlying
astrophysical and biological processes.
%
\citet{maccone_lognormals_2014} argues that this distribution should
be a log normal.

The power law and exponential statistical distributions are among the
most common patterns found in natural phenomena.
%
For example, the distribution of the frequency of words in many
languages is known to follow a Zipf's law (which is a power law).
%
This distributions arise any time a phenomena is characterized by
commonly occurring small events and rarely produced large events
\citep[e.g. ][]{adamic_zipf_2000}.
%
Zipf law also describes population ranks of cities in various
countries, corporation sizes, income rankings, ranks of number of
people watching the same TV channel.
%
The magnitudes of earthquakes, hurricanes, volcanic eruptions and
floods; the sizes of meteorites or the losses caused by business
interruptions from accidents, are also well described by power laws
\citep{sornette_critical_2006, benguigui_classificacion_2016}.
%
The power law behavior has been observed in a variety of systems,
including for example stock market fluctuations, sizes of computer
files and word frequency in languages \citep{mitzenmacher_brief_2004,
newman_power_2005, simkin_theory_2006}
%
Power laws have also been widely used in biological sciences, e.g., in
analyzing connectivity patterns in metabolic networks
\citep{jeong_large_2000} and in the number of species observed per
unit area in ecology \citep{martin_origin_2006, frank_common_2009}.
% 
More examples can be found in the literature
\citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009,
maccone_evolution_2014, maccone_lognormals_2014,
benguigui_classificacion_2016}

In this work, the exponential distribution of lifespan and waiting
times is justified by considering the hypothesis that the process of
appearance of life in the galaxy is homogeneous and stationary.
%
That is, there is no preferred location within the GHZ for the
spontaneous appearance of life, and the emergence of a \ceti{} is
independent of the existence of previous nodes in the galaxy.
%            
This is equivalent to proposing a Poisson process for the emergence of
\cetis{}, since there is a close relation between the number of events
in time or space and the waiting time or separation, respectively
\citep[e.g., ][]{ross_simulation_2012}.
%
That is, these are two alternative approaches to describing the same
process, a Poisson distribution for the number of events implies an
exponential distribution for their separations, and vice versa.
%
It should be emphasized that the exponential laws used in this work
are assumed as part of the working hypothesis, and instead of
analyzing results from a particular parameter chosen ad--hoc, we
explore the hypothesis space and analyze the impact of the values of
these parameters on the results.
%}}}

>>>>

In our simulations, we assume the simplest configuration for the
growth of the sphere of first contact. For example, we discard the
possibility of stellar colonisation
\citep[e.g.][]{newman_galactic_1981, walters_interstellar_1980,
starling_virulence_2013, barlow_galactic_2012, jeong_large_2000,
maccone_mathematical_2011}. We also assume that communication is
equally likely in all directions, i.e., we assume isotropic
communication in all cases. Different communication efficiencies or
detection methods are straightforward to carry out in the simulations
for more detailed and complex scenarios. This approach, however, is
beyond this work because it obscures the experimental setup and make
the results less clear. The assumptions we accept imply that the
results are independent of whether intelligent agents are organic or
artificial. Moreover, the causes of the limited lifetime of a
civilization can be natural (astrophysical phenomena), caused by auto
destruction or by external factors, to name a few. However, we assume
that these events are sufficiently numerous in the Galaxy so that a
statistical model is plausible. The failure of this hypothesis would
imply extreme values for the parameters that represent the density of
nodes.

@@

The temporal structure of the process is defined by two distribution parameters. One of them represent the mean time interval that a \ceti{} can emit and receive signals, and the other the mean time interval between the emergence of consecutive \cetis{}. The spatial structure of the simulation is given by the size and shape of the Galactic Habitable Zone and the maximum distance a signal can travel to be detected (D$_{max}$). The parameters for the temporal distributions also determine the spatial properties, since the density of active \cetis{} in the galaxy depend on these two parameters. For example, a small $\tau_a$ and a large $\tau_s$ will produce a galaxy densely populated with \cetis{}. Also, some hypotheses regarding the shapes of the distributions of the temporal parameters must be made in order to complete the simulation. \citet{forgan_spatiotemporal_2011} argues that the times at which different civilisations become intelligent follow a Gaussian distribution, and then the distribution of inter-arrival times is an inverse exponential. We assume that the distribution of the times of A events is a stationary Poisson process, and then the distribution of the times between the appearance of new consecutive \cetis{} is exponential. Regarding the duration of a \ceti{}, we propose that the distribution of the duration of \cetis{} is a stationary exponential distribution. There are no clear arguments to conclude a statistical law for the later distribution, so that we propose it as a working hipothesis.  A theoretical statistical distribution of the lifespan of civilisations would rely preferentially on the basis of the underlying astrophysical and biological processes \citep{balbi_impact_2018}. For example, \citet{maccone_lognormals_2014} argues that this distribution should be a log normal.

The power law and exponential statistical distributions are among the most common patterns found in natural phenomena. For example, the distribution of the frequency of words in many languages is known to follow a Zipf’s law (which is a  power law). These distributions arise any time a phenomenon is characterised by commonly occurring small events and rarely produced large events \citep[e.g. ][]{adamic_zipf_2000}. Zipf law also describes population ranks of cities in various countries, corporation sizes, income rankings, ranks of the number of people watching the same TV channel, etc. The magnitudes of earthquakes, hurricanes, volcanic eruptions and floods; the sizes of meteorites or the losses caused by business interruptions from accidents, are also well described by power laws \citep{sornette_critical_2006,  benguigui_classificacion_2016}. The power law behaviour has been observed in a variety of systems.  These include, for example, stock market fluctuations, sizes of computer files or word frequency in languages \citep{mitzenmacher_brief_2004, newman_power_2005, simkin_theory_2006}. Power laws have also been widely used in biological sciences.  Some examples are the analysis of connectivity patterns in metabolic networks \citep{jeong_large_2000} or the number of species observed per unit area in ecology \citep{martin_origin_2006, frank_common_2009}. More examples can be found in the literature \citep{martin_origin_2006, maccone_KLT_2010, barabasi_scale_2009, maccone_evolution_2014,  maccone_lognormals_2014, benguigui_classificacion_2016}.

In this work, the exponential distribution of lifespan and waiting times is justified by considering the hypothesis that the process of appearance of life in the galaxy is homogeneous and stationary. That is, there is not a preferred region within the GHZ for the spontaneous appearance of life, and the emergence of a \ceti{} is independent of the existence of previous nodes in the galaxy. This is equivalent to proposing a Poisson process for the emergence of \cetis{}, given the relation between the number of events in time or space and the waiting time or separation, respectively
\citep[e.g., ][]{ross_simulation_2012}. That is, these are two alternative approaches to describing the same process, a Poisson distribution for the number of events implies an exponential distribution for their separations, and vice versa. It should be emphasised that the exponential laws used in this work are assumed as part of the working hypothesis, and instead of analysing results from a particular parameter chosen ad hoc, we explore the hypothesis space and analyse the impact of the values of these parameters on the results.

%-------------------------------------------------------------------
%%%%%%%%%%%%%%%%%%%%%%%%%% FIRST CHECK %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%-------------------------------------------------------------------    
\rule{8cm}{12pt} \bigskip
